SPARQL on the Open, Decentralised Web

### 2016-07-24 dissertation/problems.tex
on distributions

Moreover, an ambition of the Semantic Web is to deal well with
heterogeneous data, and it has been known for long \cite{1385469} that:

\begin{problem}\label{prob:zipf}
The distribution of data is often dominated by a power-law.
\end{problem}
However, 
\begin{problem}\label{prob:uniform}
The data distribution is often unknown to the query engine, and so,
following the principle of indifference, assumes uniform
distributions.
\end{problem}

>>>>

Moreover, an ambition of the Semantic Web is to deal well with
heterogeneous data and it has been known for long \cite{1385469} that
power laws are frequently encountered when fitting a distribution to
data. As noted in the discussion of Example~\ref{ex:bgp} in
Section~\ref{sec:introinteract}, a query planner would often rely on
statistics, but:

\begin{problem}\label{prob:unknowndist}
The data distribution is often unknown to the query engine.
\end{problem}
Only simple summary statistics may be available. The discussion of
Example~\ref{ex:bgp} illustrates that even if the distribution of data
given for example a predicate is known, 
\begin{problem}\label{prob:unknownconddist}
\emph{conditional} data distributions are usually unknown to the query engine.
\end{problem}
For example, the population of geographical places may be uniformly
distributed in a given data set, but the population given that the
places are of type city may be a power law. This could have very
significant impact on the overall performance, since in the latter
case, the query planner may consider a BGP like:
\begin{example}{Skew distribution}
\begin{verbatim}
?place ex:population ?population ;
       a ex:City .
\end{verbatim}
\end{example}
as a unit.\todo{Move to preliminaries?}
